A Unified Treatment of Tractability for Approximation Problems Defined on Hilbert Spaces

### 2023-09-21 GeneralizedTractability2023August10.tex
Updates from Overleaf

\newtheorem{theorem}{Theorem}
\newdefinition{definition}{Definition}
\newdefinition{example}{Example}
\newproof{proof}{Proof}

@@

\author[1]{Fred J. Hickernell}
\ead{hickernell@iit.edu}
\author[2]{Peter Kritzer}
\ead{peter.kritzer@oeaw.ac.at}
\author[1]{Onyekachi Osisiogu}\ead{oosisiogu@iit.edu}
\date{\today}
\affiliation[1]{organization={Department of Applied Mathematics, Illinois Institute of Technology},
addressline={10 West 32${}^{\text{nd}}$ Street, Room 220},
postcode={60616},
city={Chicago, IL},
country={United States}}
\affiliation[2]{organization={Johann Radon Institute for Computational and Applied Mathematics (RICAM), Austrian Academy of Sciences},
addressline={Altenbergerstr. 69},
postcode={4040},
city={Linz},
country={Austria}}

\begin{abstract}
    A large literature specifies conditions under which the information complexity for a sequence of numerical problems defined for dimensions $1, 2, \ldots$ grows at a moderate rate, i.e., the sequence of problems is \emph{tractable}.  Here, we focus on the situation where the space of available information consists of all linear functionals and the problems are defined as operator mappings between Hilbert spaces.  We unify the proofs of known tractability results and generalize a number of existing results.  These generalizations are expressed as five theorems that provide equivalent conditions for (strong) tractability in terms of sums of functions of the singular values.
\end{abstract}

@@

\medskip

TODO: We should check for consistency of the use of the $\forall$ symbol in displayed formulas. Sometimes we use it, sometimes we do not, sometimes it is preceded by a comma, sometimes not, etc. \fred{Thanks, Kachi}

The information complexity of a problem is the number of function data required to solve the problem within the desired error tolerance.  A problem is tractable if the information complexity does not grow too large as the dimension of the problem increases or the error threshold decreases.  There are a number of tractability results for numerical problems, beginning in the 1990s with \cite{W94a,W94b}, spawning numerous articles since then, and filling a series of three impressive volumes \cite{NovWoz08a,NovWoz10a,NovWoz12a}.  Research on tractability continues at a vigorous pace.  This article aims to unify and generalize many existing tractability results for approximation problems where all linear functional data is available.

@@

The goal is to find a sequence of approximate solution operators, $\{\APP_d:\cb_d \times (0,\infty) \to \cg_d\}_{d \in \naturals}$, defined on the unit ball of functions in $\cf_d$, denoted $\cb_d$, which  satisfy an absolute error criterion:
\begin{equation}
    \label{eq:errcrit}
    \norm[\cg_d]{\SOL_d(f) - \APP_d(f,\varepsilon)} \le \varepsilon \qquad \forall f \in \cb_d, \ \varepsilon \in (0,\infty).
\end{equation}
Here, $\APP_d(f,\varepsilon)$ is allowed to depend on arbitrary linear functionals.  Note that we also allow adaptive information, i.e., the linear functionals may sequentially depend on each other.

The information complexity of the problem $\SOL_d$ is given by the function 
\[
\comp:(0,\infty) \times \naturals \to \natzero, 
\]
where $\comp(\varepsilon,d)$ is the number of linear functionals required by the best admissible algorithm to satisfy the error criterion \eqref{eq:errcrit}.
By definition, the \kachi{information }complexity is non-decreasing  as $\varepsilon$ tends to zero.  We also expect it to be non-decreasing as $d$ tends to infinity since $d$ is typically the number of variables for the functions in $\cf_d$.

For this case where arbitrary linear functionals are allowed, the optimal solution operator is known to be
\[
\APP_d(f,\varepsilon) = \sum_{i=1}^n \SOL_d(u_{i,d}) \ip[\cf_d]{f}{u_i},
\]
and the information complexity of our linear problem is
\begin{equation}\label{eq:comp_def}
\comp(\varepsilon, d) = \min \{n \in \natzero : \lambda_{n+1, d} \le \varepsilon\},
\end{equation}
see \cite{NovWoz08a}.

Strictly speaking, we are considering the \textit{absolute} error criterion in this paper.
Alternatively, one could also consider the \textit{normalized} error criterion, where the error in \eqref{eq:errcrit} is divided by the so-called \textit{initial error}, which is the error without sampling the function, or, equivalently, the operator norm of $\SOL_d$. In the latter case, it is known (see, e.g., again \cite{NovWoz08a}) that the information complexity is given by
\[
\comp^{\rm norm}(\varepsilon, d) = \min \{n \in \natzero : \lambda_{n+1, d} \le \varepsilon\, \lambda_{1, d}\}.
\]
This means that, essentially, in the tractability analysis below, all criteria would be normalized by the first eigenvalue $\lambda_{1,d}$. For the sake of brevity and clarity of notation, we restrict ourselves to the absolute error criterion here, but presumably analogous results hold for the normalized setting.

The question of tractability is one of determining how fast the information complexity increases as $\varepsilon$ tends to zero and/or as $d$ tends to infinity.  Although \eqref{eq:comp_def} has a relatively simple in form, its dependence on the ordering of the singular values means that it is not obvious what the behavior of $\comp(\varepsilon, d)$ is as $\varepsilon^{-1}$ and/or $d$ tend to infinity.

We want to bound the information complexity in terms of a simple function, $T$, of $\varepsilon^{-1}$ and $d$ and identify conditions on the singular values that are easier to verify that are equivalent to this bound on the information complexity holding. We note that generalized tractability specified by a function $T$ has been considered before in the literature. In particular, there are several papers by Gnewuch and Wo\'{z}niakowski (see \cite{GW06,GW08,GW09,GneWoz11a}), and also \cite[Chapter 8]{NovWoz08a} is devoted to this topic. To our understanding, however, these references mainly focus on the special case of tensor product problems, a restriction we do not make.  Moreover, the conditions for tractability derived there are mostly in terms of properties of the function $T$, and not in terms of sums of (functions of) the eigenvalues of a problem. Hence, the present paper an adds to what is presently known on this subject.

In this article, we investigate a \emph{generalized tractability function}, denoted by $T$:
\begin{equation} \label{eq:Tspec}
    T :(0,\infty) \times \mathbb{N} \times [0,\infty)^s \rightarrow (0,\infty).
\end{equation}
The conditions we require $T$ to satisfy are described in  Section \ref{sec:not}, however, the basic idea is that we define our approximation problem to be tractable if $\comp(\varepsilon, d) \le  C_{\vp}\, T(\varepsilon^{-1},d,\vp)$ for some constant, $C_{\vp}$, depending only on the parameter $\vp$. The parameter $\vp$ is an $s$-dimensional vector with $s \geq 1$, where $\vp \geq \vzero$ implies that each component of $\vp$ is non-negative. In practice, we expect $T(\varepsilon^{-1},d,\vp)$ to increase sub-exponentially  with increasing $\varepsilon^{-1}$ and/or $d$, however, most of our theorems do not require such an assumption.

\Cref{thm_main_strong_tract2,thm_main_tract2,thm_main_rest_strong_tract,thm_main_rest_tract} below all take the form
\begin{multline}
  \exists \,C_{\vp} > 0 \text{ with }
    \comp(\varepsilon, d) \le  C_{\vp}\, T(\varepsilon^{-1},\thed,\vp) \quad \forall (\varepsilon^{-1},d) \in \Omega \\
    \iff   \exists L_{\vp} \text{ with }
    \sup_{d \in \naturals}
     \sum_{i = \lceil L_{\vp}\, T(0,\thed,\vp) \rceil}^{\theUB(d)} \frac{1}{T(\lambda^{-1}_{i,d},\thed,\vp)}< \infty, \\
     \thed \in \{1,d\}, \ \theM(d) := \inf \{\varepsilon : (\varepsilon^{-1},d) \in \Omega \}, \
     \theUB(d) := \begin{cases}
        \min\{n \in \natzero : \lambda_{n+1,d} \le \theM(d) \}, &   \theM(d) > 0, \\
        \infty, & \theM(d) = 0.
        \end{cases}
\end{multline}
The differences in these theorems of are in the choices of $\thed$, $\Omega$, $\theM$, and $\theUB$.  Although some of these theorems are special cases of others, we prove each separately to allow the reader to become familiar with the arguments used as additional layers of complexity are added.
\Cref{thm:subhT} has a somewhat different form than above.

\peter{TODO: Maybe add a short paragraph on the notation used in the paper and on its structure? In particular, we should introduce
an ordering of the parameters $\vp$.}[\kachi{KACHI: I will take care of this}]

\kachi{[KACHI: I will write about notations used in this paper. Is there a need to change notations like $[1,\infty)$ to $\naturals$ and $[0,\infty)$ to $\naturals_0$]}

We denote the set of positive integers by $\naturals$ and and the set of non-negative integers by $\naturals_0$. The remainder of this paper is organized as follows. In Section \ref{sec:spt}, we give the definitions of the considered tractability notions, in particular (strong) tractability. We consider strong tractability when the information complexity is independent of the dimension, $d$. We then turn to tractability  when the information complexity may depend on both $d$ and $\varepsilon^{-1}$. In Section \ref{sec:examples} we provide examples of tractability functions for both the algebraic and exponential cases. We also consider other notions of tractability such as quasi-polynomial tractability, and a generalized notion of weak tractability. In the case of generalized weak tractability, we give the necessary and sufficient conditions on the notion of exponential tractability in Section \ref{sec:subh} \peter{[I think I do not quite understand what we mean by the last sentence.]}.

@@

For this definition of tractability to make sense we assume that
\begin{subequations} \label{eq:Tconditions}
\begin{equation} \label{eq:Tnondecreasing}
	T \text{ is non-decreasing in all variables,}
\end{equation}
which implies that the problem is expected to be no easier by decreasing the tolerance, $\varepsilon$, or increasing the dimension, $d$. (This is why we make $T$ a function of $\varepsilon^{-1}$ rather than of $\varepsilon$.) Furthermore, increasing $\vp$ allows for a possibly looser bound on the information complexity.  Since there are an infinite number of positive eigenvalues, it also makes sense to assume that
\begin{equation}\label{eq:Tlim}
	\lim_{\varepsilon \to 0} T(\varepsilon^{-1},d,\vp) = \infty \qquad \forall d \in \naturals, \ \vp \in [0,\infty)^s.
\end{equation}
Since $T(\cdot,d,\vp)$ is non-decreasing, we may define the following limit, which we assume will always be positive:
\begin{equation}\label{eq:T0}	T(0,d,\vp):=\lim_{\varepsilon\to\infty}T(\varepsilon^{-1},d,\vp) = \inf_\varepsilon T(\varepsilon^{-1},d,\vp) \ge T(0,1,\vzero) > 0.
\end{equation}
\end{subequations}
Finally, there is a technical assumption required for our analysis in \Cref{sec:spt,sec:restricted}.  There exists a $K_{\vp,\tau}$ depending on $\vp$ and $\tau$, but  independent of $\varepsilon$ and $d$, such that
\begin{equation} \label{eq:ptauassume}
	[T(\varepsilon^{-1},d,\vp)]^\tau \le K_{\vp,\tau} T(\varepsilon^{-1},d,\tau \vp),   \quad \forall \varepsilon \in (0,\infty), \ d \in \naturals, \ \vp\in[0,\infty)^s, \ \tau\in [1,\infty).
\end{equation}

@@

\begin{table}
    \caption{Common forms of the tractability function, $T$.}
{\small
\begin{equation*}
	\begin{array}{r@{\quad}c@{\quad}cc}
		\text{Tractability type} & T(\varepsilon^{-1},d,\vp)
		& T(0,d,\vp)\\
		\toprule
		\text{algebraic strong polynomial} & \max\{1,\varepsilon^{-p}\} & 1\\
		\text{algebraic polynomial} & \max\{1,\varepsilon^{-p}\}d^{q} & d^{q}\\
		\text{exponential strong polynomial} &  [\max\{1,\log(1 + \varepsilon^{-1}\}]^p & 1\\
		\text{exponential polynomial} &
		[\max\{1,\log(1 + \varepsilon^{-1})\}]^p  d^{q} & d^{q} \\
        \text{algebraic quasi-polynomial} &
        \exp\{p(1+\log(\max\{1,\log(1 + \varepsilon^{-1})\})(1+\log(d))\} &
        \exp\{p(1+\log(d))\}\\
        \text{exponential quasi-polynomial} &
        \exp\{p(1+\log(\max\{1,\log(1+\varepsilon^{-1})\}))(1+\log(d))\} &
        \exp\{p(1+\log(d))\}
	\end{array}
\end{equation*}}
\label{tab:commonT}
\end{table}

If \eqref{eq:tract_def} holds for some $\vp$, it clearly holds for larger $\vp$.  We are often interested in the optimal or smallest $\vp$ for which \eqref{eq:tract_def} holds.  Define the closures of the sets of parameters for which our (strong) tractability conditions hold:
\begin{equation*}
	\cp_{\textup{trct}} : = \{\vp^* : \eqref{eq:tract_def} \text{ holds }\forall \vp \in (\vp^*,\boldsymbol{\infty})\}, \qquad
	\cp_{\textup{strct}} : = \{\vp^* : \eqref{eq:strong_tract_def} \text{ holds }\forall \vp \in (\vp^*,\boldsymbol{\infty})\}.
\end{equation*}
If $\vp^* \in \cp_{\textup{(s)trct}}$, (strong) tractability may not hold for $\vp = \vp^*$, but it must hold for any $\vp$ whose components are all greater than the corresonding components of $\vp^*$.

@@

\begin{proof}
\textbf{Sufficient condition:}\newline
We make the first part of the argument in some generality so that it can be reused in the proof of Theorem \ref{thm_main_tract2} for tractability.  Fix any $\varepsilon > 0$ and any $d \in \naturals$, and let $\thed \in \naturals$ be arbitrary. %\delta replaced d'
Since the $\lambda_{i,d}$ are non-increasing in $i$, it follows that the $T(\lambda_{i,d}^{-1},\thed,\vp)$ are non-decreasing in $i$. In particular, for  any $N\in\naturals$ with $n\ge N$, we have
\begin{align*}
    \lambda_{n+1,d} \le \lambda_{n,d} \le \cdots \le \lambda_{N,d}
    & \implies \frac{1}{T(\lambda_{n+1, d}^{-1},\thed,\vp)} \le \frac{1}{T(\lambda_{n, d}^{-1},\thed,\vp)} \le \cdots \le \frac{1}{T(\lambda_{N, d}^{-1},\thed,\vp)} \\
     \implies \frac{1}{T(\lambda_{n+1, d}^{-1},\thed,\vp) }
    \le &\frac{1}{n-N+1} \sum_{i=N}^n  \frac{1}{T(\lambda_{i, d}^{-1},\thed,\vp) }
    \le \frac{1}{n-N+1} \sum_{i=N}^\infty  \frac{1}{T(\lambda_{i, d}^{-1},\thed,\vp)}.
\end{align*}
Thus, we can conclude from the previous line that
\begin{multline} \label{eq:implicate_a}
    n - N +1 \ge T(\varepsilon^{-1},\thed,\vp) \sum_{i=N}^\infty \frac{1}{T(\lambda_{i, d}^{-1},\thed,\vp)} \\
   \implies   \frac{1}{T(\lambda_{n+1, d}^{-1},\thed,\vp)} \le
   \frac{1}{n-N+1} \sum_{i=N}^\infty \frac{1}{T(\lambda_{i, d}^{-1},\thed,\vp) } \le \frac{1}{T(\varepsilon^{-1},\thed,\vp)} \\ \forall \thed, n, N \in \naturals.
\end{multline}
We note that the left side of the implication implies $n \ge N$ automatically since $n$ is an integer  $n - N +1$ is assumed to be positive.

@@

\bigskip
\noindent \textbf{Necessary condition:} \\
Suppose that we have strong
$T$-tractability as defined in \eqref{eq:strong_tract_def}. That is, for some $\vp \ge \vzero$ there exists a positive constant $C_{\vp}$ such that
\[
\comp(\varepsilon,d)\le C_{\vp}\, T(\varepsilon^{-1},1,\vp)
\qquad \forall \varepsilon > 0, \ d \in \mathbb{N}.
\]
Since the sequence of eigenvalues $\lambda_{1,d}, \lambda_{2,d}, \ldots $ is non-increasing, we have
\begin{equation}\label{eq:lambda_K_strong3}
\lambda_{\lfloor C_{\vp}\, T(\varepsilon^{-1},1,\vp)\rfloor +1,d}\le \varepsilon \qquad \forall \varepsilon > 0, \ d \in \mathbb{N}.
\end{equation}

For all positive $\varepsilon$, define
\begin{equation}\label{eq:iepsp}
i(\varepsilon,\vp):= \lfloor C_{\vp}\, T(\varepsilon^{-1},1,\vp)\rfloor +1, \quad
i(\infty,\vp) =\lfloor C_{\vp}\, T(0,1,\vp)\rfloor +1.
\end{equation}
Thus, it follows by \eqref{eq:lambda_K_strong3} that $\lambda_{i(\varepsilon,\vp),d} \le \varepsilon$.
Note furthermore that we always have
\[
i(\varepsilon,\vp)\le C_{\vp}\, T(\varepsilon^{-1},1,\vp)+1 \le C_{\vp} T(\lambda_{i(\varepsilon,\vp),d}^{-1},1,\vp)+1, \qquad \forall \varepsilon > 0, \ d \in \mathbb{N},
\]
since
$T(\cdot,1, \vp)$ is non-decreasing.

>>>>

\newtheorem{theorem}{Theorem}
\newtheorem{proposition}{Proposition}
\newdefinition{definition}{Definition}
\newdefinition{example}{Example}
\newproof{proof}{Proof}

@@

\author[1]{Onyekachi Emenike}
\ead{kachi.emenike12@gmail.com}
\author[1]{Fred J. Hickernell}
\ead{hickernell@iit.edu}
\author[2]{Peter Kritzer}
\ead{peter.kritzer@oeaw.ac.at}

\date{\today}
\affiliation[1]{organization={Department of Applied Mathematics, Illinois Institute of Technology},
addressline={10 West 32${}^{\text{nd}}$ Street, Room 220},
postcode={60616},
city={Chicago, IL},
country={United States}}
\affiliation[2]{organization={Johann Radon Institute for Computational and Applied Mathematics (RICAM), Austrian Academy of Sciences},
addressline={Altenbergerstr. 69},
postcode={4040},
city={Linz},
country={Austria}}

\begin{abstract}
    A large literature specifies conditions under which the information complexity for a sequence of numerical problems defined for dimensions $1, 2, \ldots$ grows at a moderate rate, i.e., the sequence of problems is \emph{tractable}.  Here, we focus on the situation where the space of available information consists of all linear functionals and the problems are defined as linear operator mappings between Hilbert spaces.  We unify the proofs of known tractability results and generalize a number of existing results.  These generalizations are expressed as five theorems that provide equivalent conditions for (strong) tractability in terms of sums of functions of the singular values of the solution operators.
\end{abstract}

@@

\noindent The information complexity of a problem is the number of function data required to solve the problem within the desired error tolerance.  A problem is tractable if the information complexity does not grow too large as the dimension of the problem increases or the error threshold decreases.  There are a number of tractability results for numerical problems, beginning in the 1990s with \cite{W94a,W94b}, spawning numerous articles since then, and filling a series of three impressive volumes \cite{NovWoz08a,NovWoz10a,NovWoz12a}.  Research on tractability continues at a vigorous pace.  This article aims to unify and generalize many existing tractability results for approximation problems where all linear functional data is available.

@@

The goal is to find a sequence of approximate solution operators, $\{\APP_d:\cb_d \times (0,\infty) \to \cg_d\}_{d \in \naturals}$, defined on the unit ball of functions in $\cf_d$, denoted $\cb_d$, which  satisfy an absolute error criterion:
\begin{equation}
    \label{eq:errcrit}
    \norm[\cg_d]{\SOL_d(f) - \APP_d(f,\varepsilon)} \le \varepsilon \qquad \forall f \in \cb_d, \ \varepsilon \in (0,\infty)\footnote{We note that the radius of the $\cb_d$ could be taken to be an arbitrary positive value, $R$, in which case the problem becomes equivalent to the original one with the tolerance $\varepsilon$ replaced by $\varepsilon/R$ since the problem and the optimal approximate soluion are both linear (see below).}.
\end{equation}
Here, $\APP_d(f,\varepsilon)$ is allowed to depend on arbitrary linear functionals.  Note that we also allow adaptive information, i.e., the linear functionals may sequentially depend on each other.

The information complexity of the problem $\SOL_d$ is given by the function 
\[
\comp:(0,\infty) \times \naturals \to \natzero, 
\]
where $\comp(\varepsilon,d)$ is the number of linear functionals required by the best admissible algorithm to satisfy the error criterion \eqref{eq:errcrit}.
By definition, the information complexity is non-decreasing  as $\varepsilon$ tends to zero.  We also expect it to be non-decreasing as $d$ tends to infinity since $d$ is typically the number of variables for the functions in $\cf_d$.

For this case where arbitrary linear functionals are allowed, the optimal approximate solution operator is known to be
\[
\APP_d(f,\varepsilon) = \sum_{i=1}^n \SOL_d(u_{i,d}) \ip[\cf_d]{f}{u_i},
\]
and the information complexity of our linear problem is
\begin{equation}\label{eq:comp_def}
\comp(\varepsilon, d) = \min \{n \in \natzero : \lambda_{n+1, d} \le \varepsilon\},
\end{equation}
see \cite{NovWoz08a}.

Strictly speaking, we are considering the \textit{absolute} error criterion in this paper.
Alternatively, one could also consider the \textit{normalized} error criterion, where the error in \eqref{eq:errcrit} is divided by the so-called \textit{initial error}, which is the error without sampling the function, or, equivalently, the operator norm of $\SOL_d$. In the latter case, it is known (see, e.g., again \cite{NovWoz08a}) that the information complexity is given by
\[
\comp^{\rm norm}(\varepsilon, d) = \min \{n \in \natzero : \lambda_{n+1, d} \le \varepsilon\, \lambda_{1, d}\}.
\]
This means that, essentially, in the tractability analysis below, all criteria would be normalized by the first \peter{singular value} $\lambda_{1,d}$. For the sake of brevity and clarity of notation, we restrict ourselves to the absolute error criterion here, but presumably analogous results hold for the normalized setting.

The question of tractability is one of determining how fast the information complexity increases as $\varepsilon$ tends to zero and/or as $d$ tends to infinity.  Although \eqref{eq:comp_def} has a relatively simple form, its dependence on the ordering of the singular values means that it is not obvious what the behavior of $\comp(\varepsilon, d)$ is as $\varepsilon^{-1}$ and/or $d$ tend to infinity.

We want to bound the information complexity in terms of a simple function, $T$, of $\varepsilon^{-1}$ and $d$, and identify conditions on the singular values that are easier to verify that are equivalent to this bound on the information complexity holding. We note that generalized tractability specified by a function $T$ has been considered before in the literature. In particular, there are several papers by Gnewuch and Wo\'{z}niakowski (see \cite{GW06,GW08,GW09,GneWoz11a}), and also \cite[Chapter 8]{NovWoz08a} is devoted to this topic. To our understanding, however, these references mainly focus on the special case of tensor product problems, a restriction we do not make.  Moreover, the conditions for tractability derived there focus mostly on relating the properties of the function $T$ to the different varieties of tractability, and not in terms of sums of (functions of) the \peter{singular values} of a problem. Hence, the present paper adds to what is presently known on this subject. 
In this article, we investigate a \emph{generalized tractability function}, denoted by $T$:
\begin{equation} \label{eq:Tspec}
    T :(0,\infty) \times \mathbb{N} \times [0,\infty)^s \rightarrow (0,\infty).
\end{equation}
The conditions we require $T$ to satisfy are described in  Section \ref{sec:not}, however, the basic idea is that we define our approximation problem to be tractable if $\comp(\varepsilon, d) \le  C_{\vp}\, T(\varepsilon^{-1},d,\vp)$ for some constant, $C_{\vp}$, depending only on the parameter $\vp$. The parameter $\vp$ is an $s$-dimensional vector with $s \geq 1$, where $\vp \geq \vzero$ implies that each component of $\vp$ is non-negative. In practice, we expect $T(\varepsilon^{-1},d,\vp)$ to increase sub-exponentially  with increasing $\varepsilon^{-1}$ and/or $d$, however, most of our theorems do not require such an assumption.

\Cref{thm_main_strong_tract2,thm_main_tract2,thm_main_rest_strong_tract,thm_main_rest_tract} below all take the form
\begin{multline}
  \exists \,C_{\vp} > 0 \text{ with }
    \comp(\varepsilon, d) \le  C_{\vp}\, T(\varepsilon^{-1},\thed,\vp) \quad \forall (\varepsilon^{-1},d) \in \Omega \\
    \iff   \exists L_{\vp} \peter{>0} \text{ with }
    \sup_{d \in \naturals}
     \sum_{i = \lceil L_{\vp}\, T(0,\thed,\vp) \rceil}^{\theUB(d)} \frac{1}{T(\lambda^{-1}_{i,d},\thed,\vp)}< \infty, \\
     \thed \in \{1,d\}, \ \theM(d) := \inf \{\varepsilon : (\varepsilon^{-1},d) \in \Omega \}, \
     \theUB(d) := \begin{cases}
        \min\{n \in \natzero : \lambda_{n+1,d} \le \theM(d) \}, &   \theM(d) > 0, \\
        \infty, & \theM(d) = 0.
        \end{cases}
\end{multline}
The differences in these theorems are in the choices of $\thed$ and $\Omega$. Although some of these theorems are special cases of others, we prove each separately to allow the reader to become familiar with the arguments used as additional layers of complexity are added.
\Cref{thm:subhT} has a somewhat different form than above.

In the following, we will denote the set of positive integers by $\naturals$ and the set of non-negative integers by $\naturals_0$. We write $\vzero$ to denote a vector with all components equal to zero, and $\vinfty$ to denote a vector with all components equal to $\infty$.
Furthermore, for two vectors $\vp=(p_1,\ldots,p_s)$ and $ \vp'=(p_1',\ldots,p_s')$ in $[0,\infty)^s$, 
we write $\vp\ge \vp'$ if $p_j\ge p'_j$ for all $j\in \{1,\ldots,s\}$. Furthermore, $\vp > \vp'$ 
if $p_j> p'_j$ for all $j\in \{1,\ldots,s\}$. The expression $\vp\in (\vp',\vp'')$ is to be interpreted as 
$\vp' < \vp < \vp''$ in that sense, and analogously for half-closed or closed intervals.

The remainder of this paper is organized as follows. In Section \ref{sec:spt}, we give the definitions of the considered tractability notions, in particular (strong) tractability. We consider strong tractability when there is an upper bound on the information complexity that is independent of the dimension, $d$. We then turn to tractability  when the bound on the information complexity may depend on both $d$ and $\varepsilon^{-1}$. In Section \ref{sec:examples} we provide examples of tractability functions for both the algebraic and exponential cases. We also consider other notions of tractability such as quasi-polynomial tractability, and introduce other notions of tractability. A generalized notion of weak tractability is given in Section \ref{sec:subh}. In Section \ref{sec:restricted}, we consider \peter{tractability on} a restricted domain .

@@

For this definition of tractability to make sense we assume that
\begin{subequations} \label{eq:Tconditions}
\begin{equation} \label{eq:Tnondecreasing}
	T \text{ is non-decreasing in all variables,}
\end{equation}
which implies that the problem is expected to be no easier by decreasing the tolerance, $\varepsilon$, or increasing the dimension, $d$\footnote{This is why we make $T$ a function of $\varepsilon^{-1}$ rather than of $\varepsilon$.} Furthermore, increasing $\vp$ allows for a possibly looser bound on the information complexity.  Since there are an infinite number of positive \peter{singular values}, it also makes sense to assume that
\begin{equation}\label{eq:Tlim}
	\lim_{\varepsilon \to 0} T(\varepsilon^{-1},d,\vp) = \infty \qquad \forall d \in \naturals, \ \vp \in [0,\infty)^s.
\end{equation}
Since $T(\cdot,d,\vp)$ is non-decreasing, we may define the following limit, which we assume will always be positive:
\begin{equation}\label{eq:T0}	T(0,d,\vp):=\lim_{\varepsilon\to\infty}T(\varepsilon^{-1},d,\vp) = \inf_\varepsilon T(\varepsilon^{-1},d,\vp) \ge T(0,1,\vzero) > 0.
\end{equation}
\end{subequations}
Finally, there is a technical assumption required for our analysis in \Cref{sec:spt,sec:restricted}.  There exists a $K_{\vp,\tau}$ depending on $\vp$ and $\tau$, but  independent of $\varepsilon$ and $d$, such that
\begin{equation} \label{eq:ptauassume}
	[T(\varepsilon^{-1},d,\vp)]^\tau \le K_{\vp,\tau} T(\varepsilon^{-1},d,\tau \vp)   \quad \forall \varepsilon \in (0,\infty), \ d \in \naturals, \ \vp\in[0,\infty)^s, \ \tau\in [1,\infty).
\end{equation}

@@

\begin{table}
    \caption{Common forms of the tractability function, $T$.}
{\small
\begin{equation*}
	\begin{array}{r@{\quad}c@{\quad}cc}
		\text{Tractability type} & T(\varepsilon^{-1},d,\vp)
		& T(0,d,\vp)\\
		\toprule
		%\text{algebraic strong polynomial} & \max\{1,\varepsilon^{-p}\} & 1\\
		\text{algebraic polynomial} & \max\{1,\varepsilon^{-p}\}d^{q} & d^{q}\\
		%\text{exponential strong polynomial} &  [\max\{1,\log(1 + \varepsilon^{-1}\}]^p & 1\\
		\text{exponential polynomial} &
		[\max\{1,\log(1 + \varepsilon^{-1})\}]^p  d^{q} & d^{q} \\
        \text{algebraic quasi-polynomial} &
        \exp\{p(1+\log(\max\{1,\log(1 + \varepsilon^{-1})\})(1+\log(d))\} &
        \exp\{p(1+\log(d))\}\\
        \text{exponential quasi-polynomial} &
        \exp\{p(1+\log(\max\{1,\log(1+\varepsilon^{-1})\}))(1+\log(d))\} &
        \exp\{p(1+\log(d))\}
	\end{array}
\end{equation*}}
\label{tab:commonT}
\end{table}

If \eqref{eq:tract_def} holds for some $\vp$, it clearly holds for larger $\vp$.  We are often interested in the optimal or smallest $\vp$ for which \eqref{eq:tract_def} holds.  Define the closures of the sets of parameters for which our (strong) tractability conditions hold:
\begin{equation*}
	\cp_{\textup{trct}} : = \{\vp^* : \eqref{eq:tract_def} \text{ holds }\forall \vp \in (\vp^*,\boldsymbol{\infty})\}, \qquad
	\cp_{\textup{strct}} : = \{\vp^* : \eqref{eq:strong_tract_def} \text{ holds }\forall \vp \in (\vp^*,\boldsymbol{\infty})\}.
\end{equation*}
If $\vp^* \in \cp_{\textup{(s)trct}}$, (strong) tractability may not hold for $\vp = \vp^*$, but it must hold for any $\vp$ whose components are all greater than the corresponding components of $\vp^*$.

@@

\begin{proof}
\textbf{Sufficient condition:}\newline
We make the first part of the argument in some generality so that it can be reused in the proof of Theorem \ref{thm_main_tract2} for tractability.  Fix any $\varepsilon > 0$ and any $d \in \naturals$, and let $\thed \in \naturals$ be arbitrary. %\delta replaced d'
Since the $\lambda_{i,d}$ are non-increasing in $i$, it follows that the $T(\lambda_{i,d}^{-1},\thed,\vp)$ are non-decreasing in $i$. In particular, for  any $N\in\naturals$ with $n\ge N$, we have
\begin{align*}
    \lambda_{n+1,d} \le \lambda_{n,d} \le \cdots \le \lambda_{N,d}
    & \implies \frac{1}{T(\lambda_{n+1, d}^{-1},\thed,\vp)} \le \frac{1}{T(\lambda_{n, d}^{-1},\thed,\vp)} \le \cdots \le \frac{1}{T(\lambda_{N, d}^{-1},\thed,\vp)} \\
     \implies \frac{1}{T(\lambda_{n+1, d}^{-1},\thed,\vp) }
    \le &\frac{1}{n-N+1} \sum_{i=N}^n  \frac{1}{T(\lambda_{i, d}^{-1},\thed,\vp) }
    \le \frac{1}{n-N+1} \sum_{i=N}^\infty  \frac{1}{T(\lambda_{i, d}^{-1},\thed,\vp)}.
\end{align*}
Thus, we can conclude from the previous line that
\begin{multline} \label{eq:implicate_a}
    n - N +1 \ge T(\varepsilon^{-1},\thed,\vp) \sum_{i=N}^\infty \frac{1}{T(\lambda_{i, d}^{-1},\thed,\vp)} \\
   \implies   \frac{1}{T(\lambda_{n+1, d}^{-1},\thed,\vp)} \le
   \frac{1}{n-N+1} \sum_{i=N}^\infty \frac{1}{T(\lambda_{i, d}^{-1},\thed,\vp) } \le \frac{1}{T(\varepsilon^{-1},\thed,\vp)} \\ \forall \thed, n, N \in \naturals.
\end{multline}
We note that the left side of the implication implies $n \ge N$ automatically since $n$ is an integer \peter{and} $n - N +1$ is assumed to be positive.

@@

\bigskip
\noindent \textbf{Necessary condition:} \\
Suppose that we have strong
$T$-tractability as defined in \eqref{eq:strong_tract_def}. That is, for some $\vp \ge \vzero$ there exists a positive constant $C_{\vp}$ such that
\[
\comp(\varepsilon,d)\le C_{\vp}\, T(\varepsilon^{-1},1,\vp)
\qquad \forall \varepsilon > 0, \ d \in \mathbb{N}.
\]
Since the sequence of \peter{singular values} $\lambda_{1,d}, \lambda_{2,d}, \ldots $ is non-increasing, we have
\begin{equation}\label{eq:lambda_K_strong3}
\lambda_{\lfloor C_{\vp}\, T(\varepsilon^{-1},1,\vp)\rfloor +1,d}\le \varepsilon \qquad \forall \varepsilon > 0, \ d \in \mathbb{N}.
\end{equation}

For all positive $\varepsilon$, define
\begin{equation}\label{eq:iepsp}
i(\varepsilon,\vp):= \lfloor C_{\vp}\, T(\varepsilon^{-1},1,\vp)\rfloor +1, \quad
i(\infty,\vp) =\lfloor C_{\vp}\, T(0,1,\vp)\rfloor +1.
\end{equation}
Thus, it follows by \eqref{eq:lambda_K_strong3} that $\lambda_{i(\varepsilon,\vp),d} \le \varepsilon$.
Note furthermore that we always have
\[
i(\varepsilon,\vp)\le C_{\vp}\, T(\varepsilon^{-1},1,\vp)+1 \le C_{\vp} T(\lambda_{i(\varepsilon,\vp),d}^{-1},1,\vp)+1 \qquad \forall \varepsilon > 0, \ d \in \mathbb{N},
\]
since
$T(\cdot,1, \vp)$ is non-decreasing.
